---
title: "MCsim"
author: "Sarah I. Murphy, Samantha Lau, Timothy Lott, Aljosa Trmcic"
date: "2/6/2020"
output: pdf_document
---

The authors listed above are current contributors to this file. Earlier versions were the work of Ariel Buehler and Michael Phillips. Review README.md [https://github.com/FSL-MQIP/MC-2020/blob/master/README.md] for more information.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Set-up environment
* load packages
* load functions
* set seed for reproducibilty
```{r}
## load packages
library(readr)

## load functions
source("UtilityFunctions.R")

## set seed
set.seed(1)
```

## 2. Load input files
```{r}
# (a) frequency data
frequency_file <- "Ariel_2017MC_SporeModel/InputFiles/ColdgrowATFreq_CrossSect.csv"
freq_import <- read.csv(frequency_file, stringsAsFactors = FALSE, header = TRUE)

# (b) growth parameter data
growth_file <- "Ariel_2017MC_SporeModel/InputFiles/GrowthParameters.csv" #make sure this contains growth parameters & growth model name
growth_import <-read.csv(growth_file, stringsAsFactors = FALSE)

# (c) initial count logMPN data
init_file <- "InitialCountsMPN.csv"
initialcount_import <- read.csv(init_file, stringsAsFactors = FALSE)

# (d) temperature data for each stage (Note for Sarah to fix temp_stages format)
temp_stages_file <- "temp_stages.csv"
temp_by_stage <- read.csv(temp_stages_file, stringsAsFactors = F, comment.char = "#")
```

## 3. Set-up dataframe

i. Set important parameters for running the simulation:
```{r}
# (a) n_sim
# Definition: Number of simulations to run (also referred to as the number of bulk tanks in earlier code; i.e., 1 simulation unit = 1 bulk tank of milk)
n_sim <-100    #can use any n_sim for testing code, but need to change to at least 10k when running experiments

# (b) n_units 
# Definition: Number of units per lot (currently, half gallon units) 
n_units <-10

# (c) start_day 
# Definition: First time point (in days) in unit's shelf life to simulate
start_day <- 1 

# (d) end_day 
# Definition: Last time point (in days) in unit's shelf life to simulate
end_day <- 24 
n_day <- 24 # seems like this is the end_day, but keeping code below for now in case my interpretation is wrong
```

ii. Prep vectors using defined values above
```{r}

# (a) bulk_tank
#Make a vector for "bulk tank" by repeating sequence of 1:n_sim for length of n_units * n_day
bulk_tank <- rep(seq(1, n_sim), each = n_units * n_day)

# (b) milk_unit
#Make a vector for "milk_unit" by repeating the sequence of 1:n_units by n_day * n_sim
milk_unit <- rep(seq(1, n_units), times = n_day * n_sim) 

# (c) AT
# Make an empty vector for "AT" with the length of n_sim * n_units * n_day
AT <- vector(mode="logical", n_sim * n_units * n_day) 

# (d) day
#Make a vector repeating days in shelf life that are simulated for each simulation run (i.e., each bulk tank) 
day <- rep(rep(seq(start_day, start_day+n_day-1), each = n_units), times = n_sim) 

# (e) count
#Make an empty vector for "count" with the length of n_sim * n_units * n_day
count <- vector(mode = "logical", n_sim * n_units * n_day) 
```

iii. Prep dataframe using vectors above (each vector becomes a column)
```{r}
data <- data.frame(bulk_tank, milk_unit, AT, day, count)
```

## 4. Prepare distributions using input data
a. Initial contamination: logMPN normal distribution
b. Frequency of allelic types
c. Temperature distribution for each stage 
```{r}
## (a) Initial contamination: logMPN normal distribution (note: How is the initialcount_import data used? Where did logMPN_mean and logMPN_sd come from?)
# initialcount_data = initialcount_import[,3]
# initialcountlog_data = initialcount_import[,4]

#Set mean and sd to use for normal distribution
logMPN_mean <- c(-0.7226627) 
logMPN_sd <- c(.9901429) 

## (b) Frequency of allelic types
#Here, the "distribution" of ATs are taken directly from the input file by creating a vector of the AT in each of the rows in the the column containing the AT

AT_freq <- freq_import$rpoB.allelic.type 

## (c) Temperature distribution for each stage (Note for Sarah to fix)
#Here, the distribution and parameters are currently already written into the input file
temp_by_stage

```

## 5. Sample distributions

i. Sample distributions
    a. Sample MPN_init from the initial MPN
    b. Sample AT from AT_freq
    c. Sample temp from temp_by_stage
```{r}
## (a) Sample MPN_init from the initial MPN
#Sample the logMPN distribution 
logMPN_samp <- rnorm(n_sim, logMPN_mean, logMPN_sd) 

#Convert logMPN_samp back to MPN_samp
MPN_samp <- 10^logMPN_samp 

#Convert MPN for each sample (here, unique for each simulation aka bulk tank) to equivalent in milk unit of interest (here, unit = half gallon; 1892.71 mL in half gallon)
MPN_samp_halfgal <- MPN_samp * 1900

#Sample the MPN_init distribution
MPN_init<-vector()
for (i in 1:n_sim){
  MPN_init_samp <-rep(rpois(n_units, MPN_samp_halfgal[i]), times = n_day)
  MPN_init<-c(MPN_init, MPN_init_samp)}

## (b) Sample the AT_freq distriubtion
allele <- vector()
for (i in 1:n_sim){
  allele_samp <- rep(sample(AT_freq, n_units, replace = T), times = n_day)
  allele <- c(allele, allele_samp)}

## (c) Sample the temperature distribution
temps <- vector()
for (i in 1:n_sim){
  for (j in 1:nrow(temp_by_stage)){
    stage_row <- temp_by_stage[j, ]
    n_times <- stage_row$endTime - stage_row$beginTime + 1
    params <- as.numeric(unlist(strsplit(stage_row$parameters, " ")))
    temp_mean <- params[[1]]
    temp_sd <- params[[2]]
    temp_sample <- rep(rnorm(n_units, temp_mean, temp_sd), times = n_times)
    temps <- c(temps, temp_sample)}}
```

ii. Add results from sampling distribution in (i) to dataframe
```{r}
## (a) Reformat and add initial logMPN to dataframe
#Convert MPN_init from half-gallon to mLs
MPN_init_mL <- MPN_init / 1900
#remove 0's from the data and replace with detection limit
MPN_init_mL[MPN_init_mL == 0] <- 0.01;
#Add logMPN_init to dataframe
data$logMPN_init <- log10(MPN_init_mL) 

## (b) Add AT to dataframe
data$AT<-allele 

## (c) #Add temp to dataframe
data$temp <- temps 
```

iii. Prepare empty vectors for newTemp, newMu, newLag with length n_sim * n_units * n_day
```{r}
data$newTemp <- vector(mode="logical", n_sim * n_units * n_day)
data$newMu<- vector(mode="logical", n_sim * n_units * n_day)
data$newLag <- vector(mode="logical", n_sim * n_units * n_day)
```

## 6. Calculate log10N for each row in dataframe
```{r}
##Get the AT and day from the data frame, get growth parameters depending on the AT
# Simulation   ----
for (i in 1:(n_sim *n_units * n_day)){
  #Find row in growth parameter data that corresponds to allele sample
  allele_index <- which(growth_import$rpoBAT == data$AT[i]) 
  row <- data[i, ]
  prev_row <- getPrevRow(data, row$bulk_tank, row$milk_unit, row$day)
  update <- ifelse(nrow(prev_row) == 0 || row$temp != prev_row$temp, T, F)
  #calculate the new growth parameters using the square root model and our
  #sampled temperature
  
    newT <- row$temp
    newLag <- ifelse(update, 
                     lagAtNewTemp(row$day, newT, growth_import$lag[allele_index]),
                     prev_row$newLag)
     old_lag <- ifelse(row$day==1, newLag, prev_row$newLag )
     # newLag <- ifelse(update & row$day <= old_lag & row$day > 1, 
     #                  adjustLag(row$day, old_lag, newLag),
     #                  old_lag)
     newLag <- ifelse(update & row$day > 1, 
                      adjustLag(row$day, old_lag, newLag, restartExp = F),
                      old_lag)
    newMu <-  ifelse(update, 
                     muAtNewTemp(newT, growth_import$mumax[allele_index]),
                     prev_row$newMu)
  data$newTemp[i] <- newT
  data$newLag[i] <- newLag
  data$newMu[i] <- newMu
  
  #Calculate the log10N count using our new growth parameters
  newCount <- log10N_func(row$day, newLag, newMu,data$logMPN_init[i],growth_import$LOG10Nmax[allele_index])
  oldCount <- log10N_func(row$day-1, newLag, newMu,data$logMPN_init[i],growth_import$LOG10Nmax[allele_index])
  data$count[i] <- ifelse(row$day==1, newCount,
                          prev_row$count + (newCount-oldCount))
}
```

## 7. Save results
```{r}
save(data,file="MCsim_result_df.RData")
```




