---
title: "MCsim"
author: "Sarah I. Murphy, Samantha Lau, Timothy Lott, Aljosa Trmcic"
date: "2/6/2020"
output: pdf_document
---

The authors listed above are current contributors to this file. Earlier versions were the work of Ariel Buehler and Michael Phillips. Review README.md [https://github.com/FSL-MQIP/MC-2020/blob/master/README.md] for more information.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Set-up environment
* load packages
* load functions
* set seed for reproducibilty
```{r}
## load packages
library(readr)
library(censReg)
library(fitdistrplus)
library(splitstackshape)

## load functions
source("UtilityFunctions.R")

## set seed
set.seed(1)
```

## 2. Load input files
* load
```{r}
# (a) AT/ST frequency data
# spore
spore_ATfrequency_file <- "Ariel_2017MC_SporeModel/InputFiles/ColdgrowATFreq_CrossSect.csv"
spore_ATfreq_import <- read.csv(spore_ATfrequency_file, stringsAsFactors = FALSE, header = TRUE)
# ppc
ppc_STfrequency_file <- "Sam_FFAR_PPCmodel/InputFiles/Frequency_ONLYSAMISOLATES_remove13and23_020120.csv"
ppc_STfreq_import <- read.csv(ppc_STfrequency_file, stringsAsFactors = FALSE, header = TRUE)

# (b) growth parameter data; make sure this contains growth parameters & growth model name
# spore 
spore_growth_file <- "Ariel_2017MC_SporeModel/InputFiles/GrowthParameters.csv" 
spore_growth_import <-read.csv(spore_growth_file, stringsAsFactors = FALSE)
# ppc
ppc_growth_file <- "Sam_FFAR_PPCmodel/InputFiles/GrowthParameters_remove13and23_020120.csv" 
ppc_growth_import <-read.csv(ppc_growth_file, stringsAsFactors = FALSE)

# (c) initial count logMPN data
# init_file <- "InitialCountsMPN.csv"
# initialcount_import <- read.csv(init_file, stringsAsFactors = FALSE)

# spore
spore_init_file <- read.csv("Ariel_2017MC_SporeModel/InputFiles/CrossSectional_RawMPN_3.6.17.csv")
# ppc
ppc_init_file <- read_csv("Sam_FFAR_PPCmodel/InputFiles/Initialmicrodata_MCinput_SL_021420.csv")

# (d) temperature data for each stage (Note for Sarah to fix temp_stages format)
temp_stages_file <- "temp_stages.csv"
temp_by_stage <- read.csv(temp_stages_file, stringsAsFactors = F, comment.char = "#")

# (e) PPC frequency data
source("FreqPPC.R")
```

* Check input files for spore & ppc have the same structure and then modify here as necessary to match
    * For the sake of simplifying the code, in PPC files we will temporarily call "ST", "AT" instead to match spore fies
```{r}
# modify frequency data file structure for ppc to match spore
ppc_STfreq_import <- ppc_STfreq_import %>% 
  rename(RealAT = X16S_ST)
ppc_STfreq_import$ClosestAT <- ppc_STfreq_import$RealAT
ppc_STfreq_import <- ppc_STfreq_import[c(5:6)]

ppc_growth_import <- ppc_growth_import %>% 
  rename(mumax = mu,
         LOG10N0 = N0,
         LOG10Nmax = Nmax,
         model_name = modelname)  

spore_growth_import <- spore_growth_import %>%
  .[c(1:5)]%>%
  rename(AT = rpoBAT)

ppc_init_file <- ppc_init_file %>%
  .[c(1,3)]%>%
  rename(CFU = `Initial conc`)
ppc_init_file$log10CFU <- log10(ppc_init_file$CFU)
```

## 3. Set-up dataframe

i. Set important parameters for running the simulation:
```{r}
# (a) n_sim
# Definition: Number of simulations to run (also referred to as the number of bulk tanks in earlier code; i.e., 1 simulation unit = 1 lot of milk)
n_sim <-100 #can use any n_sim for testing code, but need to change to at least 10k when running experiments

# (b) n_units 
# Definition: Number of units per lot (currently, half gallon units) 
n_units <- 10

# (c) start_day 
# Definition: First time point (in days) in unit's shelf life to simulate
start_day <- 1 

# (d) end_day 
# Definition: Last time point (in days) in unit's shelf life to simulate
end_day <- 24 
n_day <- 24 # seems like this is the end_day, but keeping code below for now in case my interpretation is wrong
```

ii. Prep vectors using defined values above
```{r}

# (a) lot_id
#Make a vector for "lot_id" by repeating sequence of 1:n_sim for length of n_units * n_day
lot_id <- rep(seq(1, n_sim), each = n_units * n_day)

# (b) milk_unit
#Make a vector for "milk_unit" by repeating the sequence of 1:n_units by n_day * n_sim
milk_unit <- rep(seq(1, n_units), times = n_day * n_sim) 

# (c) STorAT
# Make an empty vector for "STorAT" with the length of n_sim * n_units * n_day
STorAT <- vector(mode="logical", n_sim * n_units * n_day) 

# (d) day
#Make a vector repeating days in shelf life that are simulated for each simulation run (i.e., each bulk tank) 
day <- rep(rep(seq(start_day, start_day+n_day-1), each = n_units), times = n_sim) 

# (e) count
#Make an empty vector for "count" with the length of n_sim * n_units * n_day
count <- vector(mode = "logical", n_sim * n_units * n_day) 

# (f) spoilage_type
#Make an empty vector for "spoilage_type" with the length of n_sim * n_units * n_day
spoilage_type <- vector(mode = "logical", n_sim * n_units * n_day) 
```

iii. Prep dataframe using vectors above (each vector becomes a column)
```{r}
data <- data.frame(lot_id, milk_unit, STorAT, day, count,spoilage_type)
```

## 4. Prepare distributions using input data
a. Initial contamination: logMPN normal distribution
b. Frequency of allelic types
c. Temperature distribution for each stage 
```{r}
## (a) Initial contamination: logMPN normal distribution (note: How is the initialcount_import data used? Where did logMPN_mean and logMPN_sd come from?)

# determine appropriate normal distribution based on input data 
# (i) spore data
spore_init_file$log10left <- log10(spore_init_file$left)
spore_init_file$log10right <- log10(spore_init_file$right)
spore_init_file$log10MPN <- log10(spore_init_file$MPN)
cens_data <- spore_init_file[,c("log10left","log10right")]
names(cens_data) <- c("left","right")
spore_fit <- fitdistcens(censdata = cens_data,distr = "norm")
hist(spore_init_file$log10MPN,freq=F,xlim = c(-4,4))
curve(dnorm(x,mean=spore_fit$estimate[1],sd = spore_fit$estimate[2]),add=T)
plot(spore_fit)
summary(spore_fit)

spore_log10MPN_mean <- spore_fit$estimate[1]
spore_log10MPN_sd <- spore_fit$estimate[2]

# (ii) ppc data; https://stats.stackexchange.com/questions/132652/how-to-determine-which-distribution-fits-my-data-best
ppc_fit <- fitdist(data = ppc_init_file$log10CFU,distr = "norm")
plot(ppc_fit) #fit looks OK

ppc_log10CFU_mean <- ppc_fit$estimate[1]
ppc_log10CFU_sd <- ppc_fit$estimate[2]

## (b) Frequency of allelic types
#Here, the "distribution" of STs or ATs are taken directly from the input file by creating a vector of the ST or AT in each of the rows in the the column containing the ST or AT
# (i) spore data
AT_freq <- spore_ATfreq_import$ClosestAT 
# (ii) ppc data
ST_freq <- ppc_STfreq_import$ClosestAT 

## (c) Temperature distribution for each stage (Note for Sarah to fix)
#Here, the distribution and parameters are currently already written into the input file
temp_by_stage

## (d) Frequency of PPC
PPC_freq <- df_freqPPC$exceed20000_d10 #can change this to change by samplings or plants (or can rewrite code below)
```

## 5. Sample distributions

i. Sample distributions
    a. Sample MPN_init from the initial MPN
    b. Sample AT from AT_freq
    c. Sample temp from temp_by_stage
```{r}
## (a) Sample init_count from both the spore & ppc data
# (i) spore
#Sample the spore_log10MPN distribution 
spore_log10MPN_samp <- rnorm(n_sim, spore_log10MPN_mean, spore_log10MPN_sd) 

#Convert spore_log10MPN_samp back to spore_MPN_samp
spore_MPN_samp <- 10^spore_log10MPN_samp

#Convert MPN for each sample (here, unique for each simulation aka bulk tank) to equivalent in milk unit of interest (here, unit = half gallon; 1892.71 mL in half gallon)
spore_MPN_samp_halfgal <- spore_MPN_samp * 1900

#Sample the MPN_init distribution
spore_MPN_init<-vector()
for (i in 1:n_sim){
  spore_MPN_init_samp <-rep(rpois(n_units, spore_MPN_samp_halfgal[i]), times = n_day)
  spore_MPN_init<-c(spore_MPN_init, spore_MPN_init_samp)}

## (b) Sample the AT_freq distriubtion
allele <- vector()
for (i in 1:n_sim){
  allele_samp <- rep(sample(AT_freq, n_units, replace = T), times = n_day)
  allele <- c(allele, allele_samp)}

## (c) Sample the temperature distribution
temps <- vector()
for (i in 1:n_sim){
  for (j in 1:nrow(temp_by_stage)){
    stage_row <- temp_by_stage[j, ]
    n_times <- stage_row$endTime - stage_row$beginTime + 1
    params <- as.numeric(unlist(strsplit(stage_row$parameters, " ")))
    temp_mean <- params[[1]]
    temp_sd <- params[[2]]
    temp_sample <- rep(rnorm(n_units, temp_mean, temp_sd), times = n_times)
    temps <- c(temps, temp_sample)}}
```

ii. Add results from sampling distribution in (i) to dataframe
```{r}
## (a) Reformat and add initial logMPN to dataframe
#First convert MPN_init from half-gallon to mLs
MPN_init_mL <- MPN_init / 1900
#Also need to remove 0's from the data and replace with detection limit
MPN_init_mL[MPN_init_mL == 0] <- 0.01;
#Add logMPN_init to dataframe
data$logMPN_init <- log10(MPN_init_mL) 

## (b) Add AT to dataframe column STorAT
data$STorAT <- allele 

## (c) #Add temp to dataframe
data$temp <- temps 
# OR just set what the temperature should be for all:
data$temp <- 6

```

iii. Prepare empty vectors for newTemp, newMu, newLag with length n_sim * n_units * n_day
```{r}
data$newTemp <- vector(mode="logical", n_sim * n_units * n_day)
data$newMu <- vector(mode="logical", n_sim * n_units * n_day)
data$newLag <- vector(mode="logical", n_sim * n_units * n_day)
```

## 6. Calculate log10N for each row in dataframe
```{r}
##Get the ST or AT and day from the data frame, get growth parameters depending on the ST or AT
# Simulation   ----
for (i in 1:(n_sim *n_units * n_day)){
  #Find row in growth parameter data that corresponds to allele sample
  allele_index <- which(growth_import$rpoBAT == data$STorAT[i]) 
  row <- data[i, ]
  prev_row <- getPrevRow(data, row$lot_id, row$milk_unit, row$day)
  update <- ifelse(nrow(prev_row) == 0 || row$temp != prev_row$temp, T, F)
  #calculate the new growth parameters using the square root model and our
  #sampled temperature
  
    newT <- row$temp
    newLag <- ifelse(update, 
                     lagAtNewTemp(row$day, newT, growth_import$lag[allele_index]),
                     prev_row$newLag)
     old_lag <- ifelse(row$day==1, newLag, prev_row$newLag )
     # newLag <- ifelse(update & row$day <= old_lag & row$day > 1, 
     #                  adjustLag(row$day, old_lag, newLag),
     #                  old_lag)
     newLag <- ifelse(update & row$day > 1, 
                      adjustLag(row$day, old_lag, newLag, restartExp = F),
                      old_lag)
    newMu <-  ifelse(update, 
                     muAtNewTemp(newT, growth_import$mumax[allele_index]),
                     prev_row$newMu)
  data$newTemp[i] <- newT
  data$newLag[i] <- newLag
  data$newMu[i] <- newMu
  
  #Calculate the log10N count using our new growth parameters
  newCount <- log10N_func(row$day, newLag, newMu,data$logMPN_init[i],growth_import$LOG10Nmax[allele_index])
  oldCount <- log10N_func(row$day-1, newLag, newMu,data$logMPN_init[i],growth_import$LOG10Nmax[allele_index])
  data$count[i] <- ifelse(row$day==1, newCount,
                          prev_row$count + (newCount-oldCount))
}

```

## 7. Save results
```{r}
save(data,file="MCsim_result_df.RData")
```




