---
title: "MCsim_020620"
author: "Sarah I. Murphy"
date: "2/6/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set-up environment
* load packages
* load functions
* set seed for reproducibilty
```{r}
## load packages
library(readr)

## load functions
source("UtilityFunctions.R")

## set seed
set.seed(1)
```

# Load input files
```{r}
#(1) frequency data
frequency_file <- "Frequency.csv"
freq_import <- read.csv(frequency_file, stringsAsFactors = FALSE, header = TRUE)

#(2) growth parameter data
growth_file <- "GrowthParameters.csv" #make sure this contains growth parameters & growth model name
growth_import <-read.csv(growth_file, stringsAsFactors = FALSE)

#(3) initial count logMPN data
init_file <- "InitialCountsMPN.csv"
initialcount_import <- read.csv(init_file, stringsAsFactors = FALSE)

#(4) temperature data for each stage
temp_by_stage <- read.csv(temp_stages_file, stringsAsFactors = F, comment.char = "#")
```

# Set-up dataframe
* Set important parameters for running the simulation
* Prep vectors using the defined parameters
* Prep dataframe using the vectors
```{r}
##Set important parameters for running the simulation:

# (1) n_sim
# Definition: Number of simulations to run (also referred to as the number of bulk tanks in earlier code; i.e., 1 simulation unit = 1 bulk tank of milk)
n_sim <-100    #1000 is for testing and exploring, experiments require at least 10k

#(2) n_units 
# Definition: Number of units per lot (currently, half gallon units) 
n_units <-10

# (3) start_day 
# Definition: First time point (in days) in unit's shelf life to simulate
start_day <- 1 

# (4) end_day 
# Definition: Last time point (in days) in unit's shelf life to simulate
end_day <- 24 
n_day <- 24 # seems like this is the end_day, but keeping code below for now in case my interpretation is wrong

## Prep vectors using defined values above
# (1) bulk_tank
#Make a vector for "bulk tank" by repeating sequence of 1:n_sim for length of n_units * n_day
bulk_tank <- rep(seq(1, n_sim), each = n_units * n_day)

# (2) milk_unit
#Make a vector for "milk_unit" by repeating the sequence of 1:n_units by n_day * n_sim
milk_unit <- rep(seq(1, n_units), times = n_day * n_sim) 

# (3) AT
# Make an empty vector for "AT" with the length of n_sim * n_units * n_day
AT <- vector(mode="logical", n_sim * n_units * n_day) 

# (4) day
#Make a vector repeating days in shelf life that are simulated for each simulation run (i.e., each bulk tank) 
day <- rep(rep(seq(start_day, start_day+n_day-1), each = n_units), times = n_sim) 

# (5) count
#Make an empty vector for "count" with the length of n_sim * n_units * n_day
count <- vector(mode = "logical", n_sim * n_units * n_day) 

## Prep dataframe using vectors above (each vector becomes a column)
data <- data.frame(bulk_tank, milk_unit, AT, day, count)
```

## Prep distributions to sample from using input data
```{r}
## How is the initialcount_import data used? Prep distibution for logMPN??
# initialcount_data = initialcount_import[,3]
# initialcountlog_data = initialcount_import[,4]

## Prep values needed for logMPN normal distribution
logMPN_mean <- c(-0.7226627) #where does this come from?
logMPN_sd <- c(.9901429) #where does this come from?

## Sample from logMPN normal distribution
logMPN_samp = rnorm(n_sim, logMPN_mean, logMPN_sd)

## Convert logMPN_samp back to MPN
MPN_samp = 10^logMPN_samp

## Convert MPN for each sample (here, unique for each simulation aka bulk tank) to equivalent in milk unit of interest (here, unit = half gallon)
MPN_samp_halfgal = MPN_samp * 1900 #MPN per half gallon (1892.71 mL in half gallon)

```

## Calculate samples used in the monte carlo
+ Now sample the MPN distributions and the temperature distribution
```{r}


#Generate initial MPN for each half gallon from Poisson distribution
#Also sample AT for each half gallon
MPN_init<-vector()
allele <- vector()
temps <- vector()
for (i in 1:n_sim){
  MPN_init_samp <-rep(rpois(n_halfgal, MPN_samp_halfgal[i]), times = n_day)
  MPN_init<-c(MPN_init, MPN_init_samp)
  allele_samp <- rep(sample(freq_data, n_halfgal, replace = T), times = n_day)
  allele <- c(allele, allele_samp)
  #now calculate temp
  for (j in 1:nrow(stages)){
    stage_row <- stages[j, ]
    n_times <- stage_row$endTime - stage_row$beginTime + 1
    params <- as.numeric(unlist(strsplit(stage_row$parameters, " ")))
    temp_mean <- params[[1]]
    temp_sd <- params[[2]]
    temp_sample <- rep(rnorm(n_halfgal, temp_mean, temp_sd), times = n_times)
    temps <- c(temps, temp_sample)
  }
}
#add in temperature
data$temp <- temps

#Convert MPN_init from half-gallon to mLs
MPN_init_mL <- MPN_init / 1900
#remove 0's from the data and replace with detection limit
MPN_init_mL[MPN_init_mL == 0] <- 0.01;

#Now we add in those calculations to our original dataframe
data$logMPN_init <- log10(MPN_init_mL) #Add initial logMPN to data frame
data$AT<-allele #Add in AT data

data$newTemp <- vector(mode="logical", n_sim * n_halfgal * n_day)
data$newMu<- vector(mode="logical", n_sim * n_halfgal * n_day)
data$newLag <- vector(mode="logical", n_sim * n_halfgal * n_day)
```


