---
title: "UtilityFunctions"
author: "Sarah I. Murphy"
date: "2/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Utility Functions

+ muAtNewTemp
    + Purpose: Calculate the new mu parameter at new temperature.
    + Params: 
        + newTemp: the new temperature for which we calculate mu
        + oldMu: the previous mu value to adjust
        + oldTemp: the temperature corresponding to previous mu
        + T0: parameter used to calculate new mu
        + This uses the Ratkowsky square model which describes the effect of temperature on the growth of
          microorganisms
        + The paper it comes from: https://www.ncbi.nlm.nih.gov/pubmed/22417595
        + Sam checked equation and it is correct
        + The T0 is estimate dto be -3.62C based on growth curves of Paenibacillus ordorifer obtained at 4, 7, and 32C in BHI broth (N.H. Martin unpublished data)
```{r}
muAtNewTemp <- function(newTemp, oldMu, oldTemp = 6, T0 = -3.62) {
  numerator <- newTemp - T0
  denom <- oldTemp - T0
  newMu <- ((numerator / denom)^2) * oldMu
  
  return(newMu)
}
```

+ adjustLag
    + Purpose: Adjust the lag phase based on the Zwietering 1994 paper.
    + Params:  
        + t: the current timestep
        + oldLag: the lag time at the previous temperature
        + newLag: the lag time at the current temparature.
        + restartExp: If true then lag phase restarts even if already in exponential growth phase.
        + adjustmentConstant: The amount to adjust lag, the paper recommends 0.25
```{r}
adjustLag <- function (t, oldLag, newLag, restartExp = T, adjustmentConstant = 0.25) {
  #determine the amount of lag phase completed
  remainingLag <- 1 - (t / oldLag)
  if(restartExp) {
  remainingLag <- ifelse(remainingLag < 0, 0, remainingLag)
  }
  else {
  adjustedLag <- ifelse(remainingLag <=0, oldLag,
                        t + remainingLag * newLag + adjustmentConstant*newLag)
  }
  
  adjustedLag <- t + remainingLag*newLag + adjustmentConstant*newLag
  return(adjustedLag)
}
```

+ lagAtNewTemp
    + Purpose: Calculate the new lag parameter at new temperature.
    + Params:
        + newTemp: the new temperature for which we calculate lag
        + oldLag: the previous lag value to adjust
        + oldTemp: the temperature corresponding to previous lag
        + T0: parameter used to calculate new lag
```{r}
lagAtNewTemp <- function (t, newTemp, oldLag, oldTemp = 6, T0 = -3.62) {
  numerator <- oldTemp -T0
  denom <- newTemp - T0
  newLag <- ( (numerator / denom)^2) * oldLag
  return(newLag)
}

getPrevRow <- function(df, sim_run, half_gallon, day) {
  old_temp <- df[df$BT == sim_run & df$half_gal == half_gallon & df$day==day-1,] 
}
```


+ buchanan_log10N
    + Purpose:
    + Params:

+ Growth Models
  + All equations are copied from the Nlms package in R
  + URL: https://rdrr.io/cran/nlsMicrobio/src/R/growthmodels.R
  
```{r}
buchanan_log10N = function(t,lag,mumax,LOG10N0,LOG10Nmax){
  ans <- LOG10N0 + (t >= lag) * (t <= (lag + (LOG10Nmax - LOG10N0) * log(10)/mumax)) * mumax * (t - lag)/log(10) + (t >= lag) * (t > (lag + (LOG10Nmax - LOG10N0) * log(10)/mumax)) * (LOG10Nmax -     LOG10N0)
  return(ans)
}
```

+ gompertz_log10N
    + Purpose:
    + Params:
```{r}
gompertz_log10N = function(t,lag,mumax,LOG10N0,LOG10Nmax) {
  ans <- LOG10N0 + (LOG10Nmax - LOG10N0) * exp(-exp(mumax * exp(1) * (lag - t)/((LOG10Nmax - LOG10N0) * log(10)) + 1))
  return(ans)
}
```

+ baranyi_log10N
    + Purpose:
    + Params:
```{r}
baranyi_log10N = function(t,lag,mumax,LOG10N0,LOG10Nmax) {
  ans <- LOG10Nmax + log10((-1 + exp(mumax * lag) + exp(mumax * t))/(exp(mumax * t) - 1 + exp(mumax * lag) * 10^(LOG10Nmax - LOG10N0)))
  return(ans)
}
```

+ log10N wrapper function 
    + Purpose: This implements the growth model
    + You need to import a CSV file with the best fitting model pre-selected
    + Your CSV file should have a column that has a column name, model_name and filled out with buchanan, baranyi, or gompertz
```{r}
log10N_func <- function(t, lag, mumax, LOG10N0, LOG10Nmax, model_name="buchanan") {
  if (model_name == "buchanan") {
    return(buchanan_log10N(t, lag, mumax, LOG10N0, LOG10Nmax) )
  }
  else if(model_name == 'baranyi') {
    return(baranyi_log10N(t, lag, mumax, LOG10N0, LOG10Nmax) )
  }
  else if(model_name == 'gompertz') {
    return(gompertz_log10N(t, lag, mumax, LOG10N0, LOG10Nmax) )
  }
  else {
    stop(paste0(model_name, " is not a valid model name. Must be one of buchanan, baranyi, gompertz"))
  }
}
```