---
title: "MCsim_RachelFIVSModel"
author: "Rachel Evanowski, Sarah I. Murphy"
date: "8/7/2020"
output: pdf_document
---

This is a simplified version (deleted unnecessary stuff) of MCsim.Rmd found in MC-2020 on GitHub.

The authors listed above are current contributors to this file. Additional authors of MCsim.Rmd include Samantha Lau, Timothy Lott, and Aljosa Trmcic. Earlier versions were the work of Ariel Buehler and Michael Phillips. Review README.md [https://github.com/FSL-MQIP/MC-2020/blob/master/README.md] for more information.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Set-up environment
* load packages
* load functions
* set seed for reproducibilty
```{r}
## load packages
library(readr)
library(tidyverse)
library(censReg)
library(fitdistrplus)
library(splitstackshape)
library(rmutil)

## load functions
source("UtilityFunctions.R")
```

```{r knitr_options , include=FALSE}
## set seed
set.seed(1)
```
## 2. Load input files
* load
```{r}
# (a) AT/ST frequency data (from Ariel_2017MC_SporeModel on MC-2020 on GitHub)
# spore
spore_ATfrequency_file <- "InputFiles/ColdgrowATFreq_CrossSect.csv"
spore_ATfreq_import <- read.csv(spore_ATfrequency_file, stringsAsFactors = FALSE, header = TRUE)

# (b) growth parameter data; make sure this contains growth parameters & growth model name
# spore (from Ariel_2017MC_SporeModel on MC-2020 on GitHub)
spore_growth_file <- "InputFiles/GrowthParameters.csv" 
spore_growth_import <-read.csv(spore_growth_file, stringsAsFactors = FALSE)

# (c) initial microbial count data (from Ariel_2017MC_SporeModel on MC-2020 on GitHub)
# spore
spore_init_file <- read.csv("InputFiles/CrossSectional_RawMPN_3.6.17.csv")
```

* Modify naming in input files to be consistent across all MC-2020 model files
```{r}
spore_ATfreq_import$temp <- "AT"
spore_ATfreq_import$ClosestAT <- paste(spore_ATfreq_import$temp,spore_ATfreq_import$ClosestAT,sep="_")

spore_growth_import <- spore_growth_import %>%
  .[c(1:5)]%>%
  rename(STorAT = rpoBAT)
spore_growth_import$model_name <- "buchanan" #may change later
spore_growth_import$temp <- "AT"
spore_growth_import$STorAT <- paste(spore_growth_import$temp,spore_growth_import$STorAT,sep="_")
```

## 3. Set-up dataframe

i. Set important parameters for running the simulation:
```{r}
# (a) n_sim
# Definition: Number of simulations to run (also referred to as the number of bulk tanks in earlier code; i.e., 1 simulation unit = 1 lot of milk)
n_sim <- 10000 #can use any n_sim for testing code, but need to change to at least 10k when running experiments

# (b) n_units 
# Definition: Number of units per lot (currently, half gallon units) 
n_units <- 10

# (c) start_day 
# Definition: First time point (in days) in unit's shelf life to simulate
start_day <- 1 

# (d) end_day 
# Definition: Last time point (in days) in unit's shelf life to simulate
end_day <- 35 
n_day <- 35 # seems like this is the end_day, but keeping code below for now in case my interpretation is wrong
```

ii. Prep vectors using defined values above
```{r}
# (a) lot_id
#Make a vector for "lot_id" by repeating sequence of 1:n_sim for length of n_units * n_day
lot_id <- rep(seq(1, n_sim), each = n_units * n_day)

# (b) milk_unit
#Make a vector for "milk_unit" by repeating the sequence of 1:n_units by n_day * n_sim
milk_unit <- rep(seq(1, n_units), times = n_day * n_sim) 

# (c) STorAT
# Make an empty vector for "STorAT" with the length of n_sim * n_units * n_day
STorAT <- vector(mode="logical", n_sim * n_units * n_day) 

# (d) day
#Make a vector repeating days in shelf life that are simulated for each simulation run (i.e., each bulk tank) 
day <- rep(rep(seq(start_day, start_day+n_day-1), each = n_units), times = n_sim) 

# (e) count
#Make an empty vector for "count" with the length of n_sim * n_units * n_day
count <- vector(mode = "logical", n_sim * n_units * n_day) 

# (f) spoilage_type
#Make an empty vector for "spoilage_type" with the length of n_sim * n_units * n_day
spoilage_type <- vector(mode = "logical", n_sim * n_units * n_day) 
```

iii. Prep dataframe using vectors above (each vector becomes a column)
```{r}
data <- data.frame(lot_id, milk_unit, STorAT, spoilage_type, day, count)
```

## 4. Prepare distributions using input data
a. Initial contamination: logMPN normal distribution
b. Frequency of allelic types
c. Temperature distribution for each stage 
```{r}
## (a) Initial contamination: logMPN normal distribution
spore_init_file$log10left <- log10(spore_init_file$left)
spore_init_file$log10right <- log10(spore_init_file$right)
spore_init_file$log10MPN <- log10(spore_init_file$MPN)
cens_data <- spore_init_file[,c("log10left","log10right")]
names(cens_data) <- c("left","right")
spore_fit <- fitdistcens(censdata = cens_data,distr = "norm")
hist(spore_init_file$log10MPN,freq=F,xlim = c(-4,4))
curve(dnorm(x,mean=spore_fit$estimate[1],sd = spore_fit$estimate[2]),add=T)
plot(spore_fit)
summary(spore_fit)

spore_log10MPN_mean <- spore_fit$estimate[1]
spore_log10MPN_sd <- spore_fit$estimate[2]

## (b) Frequency of allelic types
#Here, the "distribution" of ATs are taken directly from the input file by creating a vector of the AT in each of the rows in the the column containing the AT
AT_freq <- spore_ATfreq_import$ClosestAT 

# (c) See below for temps

```

## 5. Sample distributions
i. Sample distributions
    a. Sample MPN_init from the initial MPN
    b. Sample AT from AT_freq
    c. Sample temp from Laplace distribution
    
```{r}
# (a) Sample MPN_init from the initial MPN
#Sample the spore_log10MPN distribution 
spore_log10MPN_samp <- rnorm(n_sim, spore_log10MPN_mean, spore_log10MPN_sd) 

#Convert spore_log10MPN_samp back to spore_MPN_samp
spore_MPN_samp <- 10^spore_log10MPN_samp

#Convert MPN for each sample (here, unique for each simulation aka lot) to equivalent in milk unit of interest (here, unit = half gallon; 1892.71 mL in half gallon)
spore_MPN_samp_halfgal <- spore_MPN_samp * 1900

#Sample the MPN_init distribution
spore_MPN_init<-vector()
for (i in 1:n_sim){
  spore_MPN_init_samp <-rep(rpois(n_units, spore_MPN_samp_halfgal[i]), times = n_day)
  spore_MPN_init<-c(spore_MPN_init, spore_MPN_init_samp)}

# (b) Sample AT from AT_freq
AT <- vector()
for (i in 1:n_sim){
 AT_samp <- rep(sample(AT_freq, n_units, replace = T), times = n_day)
  AT <- c(AT, AT_samp)}

## (c) #Add temp to dataframe
# (i) Sample the temperature distribution 
##Old version
# temps <- vector()
# for (i in 1:n_sim){ 
#   temp_samp <- rep(rlaplace(m=4.06, s=2.31), times = n_day*n_units)
#                    temps <- c(temps,temp_samp)} #https://doi.org/10.4315/0362-028X-73.2.312 (source)

temps <- rep(NA, n_sim)

for (i in 1:(n_sim)){
  number <- rlaplace(1,m=4.06,s=2.31)
  while (number > 15 | number < -1) {
    number <- rlaplace(1,m=4.06,s=2.31) #make sure that this cannot be >15 or < -1
  }
  temps[i] <- number
}

temps_v2 <- rep(temps, each = n_day*n_units) # expand temps so that repeat for n_units over n_day

# (ii) Set constant temperature (Note: Rachel, ask Sarah if want to sample from temperature distribution of consumer refrigerators, we did this for the Buffalo MC model)
# data$temp <- 6

```
ii. Apply intervention and add results from sampling distribution in (i) to dataframe
```{r}
## (a) Reformat and add initial log10MPN to dataframe
#First convert spore_MPN_init from half-gallon to mLs
spore_MPN_init_mL <- spore_MPN_init / 1900

#Apply intervention in unit of MPN/mL
data$spore_MPN_init_mL <- spore_MPN_init_mL #No intervention OR if running intervention in unit of log10MPN/mL
# data$spore_MPN_init_mL <- spore_MPN_init_mL - 0.3066 #Subtract the reduction in CFU for towel intervention
# data$spore_MPN_init_mL <- spore_MPN_init_mL - 0.0183 #Subtract the reduction in CFU for training intervention

#Also need to remove 0's from the data and replace with detection limit
data$spore_MPN_init_mL[data$spore_MPN_init_mL<=0 ] <- 0
data$spore_MPN_init_mL[data$spore_MPN_init_mL == 0] <- 0.01;

#Add spore_log10MPN_init to dataframe
data$spore_log10MPN_init_mL <- log10(data$spore_MPN_init_mL) 

#Apply intervention in unit of log10MPN/mL
# data$spore_log10MPN_init_mL <- data$spore_log10MPN_init_mL #No intervention
data$spore_log10MPN_init_mL <- data$spore_log10MPN_init_mL - 0.22 #Subtract the reduction in log10CFU for intervention

## (b) Add AT to dataframe column STorAT
data$STorAT <- AT

## (c) #Add temp to dataframe
data$newT <- temps_v2
```

## 6. Calculate log10N for each row in dataframe
```{r}
# Sarah updated this code 8/7/20 based on code that also write for Sam's PPC model;

##After troublshooting, saw that here, don't define "t" so remove from function call, otherwise get error
# 
# lagAtNewTemp <- function (t, newTemp, oldLag, oldTemp = 6, T0 = -3.62) {
#   numerator <- oldTemp -T0
#   denom <- newTemp - T0
#   newLag <- ( (numerator / denom)^2) * oldLag
#   return(newLag)
# }

# removed "t" from lagAtNewTemp
lagAtNewTemp <- function (newTemp, oldLag, oldTemp = 6, T0 = -3.62) {
  numerator <- oldTemp -T0
  denom <- newTemp - T0
  newLag <- ( (numerator / denom)^2) * oldLag
  return(newLag)
}

#make a temporary df with all except day
df_temp <- data[c(1:3,9)]
df_temp %>%
  unique() -> df_temp

# just determine newLag and newMu here
# for (i in 1:(n_sim*n_units)){
for (i in 1:(n_sim*n_units)){
  #Find row in growth parameter data that corresponds to allele sample
 allele_index <-
    which(spore_growth_import$STorAT == df_temp$STorAT[i])
  
  #calculate the new growth parameters using the square root model and our
  #sampled temperature
  newLag <- lagAtNewTemp(df_temp$newT[i], spore_growth_import$lag[allele_index])
  newMu <-  muAtNewTemp(df_temp$newT[i], spore_growth_import$mumax[allele_index])

  df_temp$newLag[i] <- newLag
  df_temp$newMu[i] <- newMu

}

####
# make a temporary df for merging
df_temp2 <- data[c(1:3,5:8)]
#merge
data2 <- merge(df_temp2,df_temp,by=c("lot_id","milk_unit","STorAT"))

#now determine count
for (i in 1:(n_sim *n_units * n_day)){
  #Find row in growth parameter data that corresponds to allele sample
  allele_index <- which(spore_growth_import$STorAT == data2$STorAT[i]) 
  
  #Calculate the log10N count using our new growth parameters
  #multiply log10N(data$day[i] by 24 since I did it in hours, not days, and ariel did it in days`
  data2$count[i] <- log10N_func(data2$day[i], data2$newLag[i],data2$newMu[i],data2$spore_log10MPN_init_mL[i],spore_growth_import$LOG10Nmax[allele_index])

}

data_intervention <- data2
save(data_intervention, file = "data_intervention.RData")
```

```{r}
## Old code as of 8/7/20; Rachel needs to re-run analysis using updated code

# for (i in 1:(n_sim * n_units * n_day)) {
#   allele_index <-
#     which(spore_growth_import$STorAT == data$STorAT[i]) #Find row in growth parameter data that corresponds to allele sample
#   row <- data[i,]
#   prev_row <- getPrevRow(data, row$lot_id, row$milk_unit, row$day)
#   update <-
#     ifelse(nrow(prev_row) == 0 || row$temp != prev_row$temp, T, F)
#   data$count[i] <-
#     log10N_func(
#       data$day[i],
#       spore_growth_import$lag[allele_index],
#       spore_growth_import$mumax[allele_index],
#       data$spore_log10MPN_init_mL[i],
#       spore_growth_import$LOG10Nmax[allele_index])}
```

## 7. Save results
```{r}
data_no_intervention <- data2
# data_intervention <- data2

save(data_no_intervention, file = "data_no_intervention.RData")
# save(data_intervention, file = "data_intervention.RData")

```




